---
title: 'JHU Data Science Specialization: Reproducible Research'
author: "Devara Izaz Fathan"
date: "8/18/2020"
output: html_document
---

# 1. Synopsis
The aim of this project is to explore the NOAA Storm Database and explore the effects of severe weather events on both population and economy.The database covers the time period between 1950 and November 2011. This database tracks characteristics of major storms and weather events in the United States, including when and where they occur, as well as estimates of any fatalities, injuries, and property damage.
The following analysis investigates which types of severe weather events are most harmful according to the following questions:

1. Across the United States, which types of events are most harmful with respect to population health?
2. Across the United States, which types of events have the greatest economic consequences?

# 2. Data Processing
## 2.1 Data Loading
Download the data from [Storm Data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2) and unzip the file using the function in R. We also have to add the library that will be used in this project.

```{r, echo=TRUE}
library(tidyverse)
library(dplyr)
library(knitr)
library(ggplot2)
url <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2'
dest <- download.file(url, 'E:/BELAJAR CODING/R/COURSERA/U.S.-National-Oceanic-and-Atmospheric-Administration-s-NOAA/repdata_data_StormData.csv.bz2')
storm_data <- read.csv('repdata_data_StormData.csv.bz2')
```

## 2.2 List of Column Names

```{r, echo=TRUE}
colnames(storm_data)
```

## 2.3 Changing the COlumn Names 
We need to change some column names that most frequently used such as PROPDMG, PROPDMGEXP, CROPDMG, CROPDMGEXP

```{r, echo=TRUE}
colnames(storm_data)[25:28] <- c('Damage', 'Exp', 'CropDamage', 'CropExp')
head(storm_data[,25:28])
```

## 2.4 Choosing Columns
We need to subset the data into table that contain column that we only use in this analysis

```{r, echo=TRUE}
storm_data <- storm_data %>% select(EVTYPE, FATALITIES, INJURIES, Damage, Exp, CropDamage, CropExp) 
head(storm_data)
```

## 2.5 Convert Exp and CropExp to Actual Value
We need to do this for easier calculation of getting actual cost.

```{r, echo=TRUE}
for(i in 1:nrow(storm_data)){
      exp <- storm_data$Exp[i]
      if(exp==''){storm_data$Exp[i] <- 0}
      if(exp=='-'){storm_data$Exp[i] <- 10^0}
      if(exp=='+'){storm_data$Exp[i] <- 10^0}
      if(exp=='0'){storm_data$Exp[i] <- 10^0}
      for (n in 1:9){
            if(exp == as.character(n)){storm_data$Exp[i] <- 10^n}
      }
      if(exp=='H'){storm_data$Exp[i] <- 10^2}
      if(exp=='K'){storm_data$Exp[i] <- 10^3}
      if(exp=='M'){storm_data$Exp[i] <- 10^6}
      if(exp=='B'){storm_data$Exp[i] <- 10^9}
      if(exp=='h'){storm_data$Exp[i] <- 10^2}
      if(exp=='m'){storm_data$Exp[i] <- 10^6}
      if(exp=='?'){storm_data$Exp[i] <- 10^0}
}
```

```{r}
for (i in 1:nrow(storm_data)){
      expcrop <- storm_data$CropExp[i]
      if(expcrop==''){storm_data$CropExp[i] <- 0}
      if(expcrop=='?'){storm_data$CropExp[i] <- 10^0}
      if(expcrop=='0'){storm_data$CropExp[i] <- 10^0}
      if(expcrop=='2'){storm_data$CropExp[i] <- 10^2}
      if(expcrop=='B'){storm_data$CropExp[i] <- 10^9}
      if(expcrop=='M'){storm_data$CropExp[i] <- 10^6}
      if(expcrop=='m'){storm_data$CropExp[i] <- 10^6}
      if(expcrop=='K'){storm_data$CropExp[i] <- 10^3}
      if(expcrop=='k'){storm_data$CropExp[i] <- 10^3}
}
```

## 2.6 Calculating Total Property and Crop Cost Damage

```{r}
storm_data$Exp <- as.numeric(storm_data$Exp)
storm_data$CropExp <- as.numeric(storm_data$CropExp)
damage <- storm_data %>% group_by(EVTYPE) %>% 
      summarize(prop_damage= sum(Damage*Exp), crop_damage = sum(CropDamage*CropExp), total_damage = (prop_damage + crop_damage)) %>%
      arrange(desc(total_damage))
```

## 2.7 Calculating Total Fatalities and Injuries

```{r, echo=FALSE}
fatality <- storm_data %>% group_by(EVTYPE)  %>% summarize(fatalities = sum(FATALITIES), injury=sum(INJURIES), total = fatalities + injury) %>% arrange(desc(total))
```
# 3. Result
## 3.1 Most Dangerous Event to Population Health

```{r}
head(fatality, 10)
```
The most harmful event to Population Health is **TORNADO** .We can see it in the histogram below.
```{r}
fatality[1:10,]  %>% ggplot(aes(EVTYPE, total)) +
      geom_bar(stat='identity') +
      xlab('Events') +
      ylab('Total Injuries and Fatalities') +
      coord_flip()
```

## 3.2 Most Dangerous Event to Economic Consequences

```{r}
head(damage, 10)
```
The most harmful Event to Economic Consequences is **FLOOD**. We can see it in the histogram below.

```{r}
damage[1:10, ] %>% ggplot(aes(EVTYPE, total_damage)) +
      geom_histogram(stat='identity') +
      xlab('Events') +
      ylab('Total Damage Cost for Property and Crop ($)') +
      coord_flip()
      
```







































